---
title: "analysis"
output: html_document
date: "2025-01-01"
---

website origin data 
https://www.kaggle.com/datasets/spscientist/students-performance-in-exams/code

resource of checking normality
https://www.sthda.com/english/wiki/normality-test-in-r

resource on linear regression
https://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/

resource on basic plots
https://environmentalcomputing.net/graphics/ggplot/ggplot-basics/


*This is fictional data*

Random notes and thoughts

Using if_else changes the TRUE/FALSE to numerical if T/F == 1/0
should literally put "0"/ "1". should change as.factor() afterwards


```{r}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable,    # converting tables to HTML
  ggpubr,        # easy ggplot2-based data visualization
  corrplot      # make correlation matrix
)
```

importing data

```{r import data}
exams <- import("StudentsPerformance.csv")
```

```{r}
skim(exams)
```

Looks like we have 5 categorical columns and 3 continuous

Look like no missing data either :)

TO DO: clean the categorical data: make one-hot-encoding

First let's clean the column names

```{r}
exams <- exams %>% janitor::clean_names()
```

```{r}
colnames(exams)
```

Checking unique values before re-coding

```{r}
unique(exams$gender)
```

```{r}
exams$gender <- as.factor(if_else(exams$gender=="male",1,0))
```

checking values in race_ethnicity

```{r}
unique(exams$race_ethnicity)
```


According to Data Dictionary on National Codes for ethnicity the groups signify the following;
Group A - White - British
Group B - White - Irish
Group C - White - Any other White background
Group D - Mixed - White and Black Caribbean
Group E - Mixed - White and Black African

```{r}
exams <- exams %>% 
  mutate(race_ethnicity_f = case_when(
    race_ethnicity == "group A" ~ 0,
    race_ethnicity == "group B" ~ 1,
    race_ethnicity == "group C" ~ 2,
    race_ethnicity == "group D" ~ 3,
    race_ethnicity == "group E" ~ 4,
  ))
```

changing into factor, just in case there is an order(?)

```{r}
exams$race_ethnicity_f <- as.factor(exams$race_ethnicity_f)
```


```{r}
unique(exams$parental_level_of_education)
```


PERHAPS CHANGES THIS LOOK INTO R4EPI LESSON!

*the past code was causing error, know it is fixed
```{r}
exams <- exams %>% 
  mutate(parental_level_of_education_f = case_when(
    parental_level_of_education == "some high school"   ~ 0,
    parental_level_of_education == "high school"        ~ 1,
    parental_level_of_education == "some college"       ~ 2,
    parental_level_of_education == "associate's degree" ~ 3,
    parental_level_of_education == "bachelor's degree"  ~ 4,
    parental_level_of_education == "master's degree"    ~ 5
  ))
```

```{r}
unique(exams$lunch)
```

```{r}
unique(exams$test_preparation_course)
```


Re-label and organizing is complete let's so more analysis

```{r}
skim(exams)
```

Let's see if the 3 scores are normally distributed: math, reading, writing

ggdensity(my_data$len, 
          main = "Density plot of tooth length",
          xlab = "Tooth length")



```{r}
ggdensity(exams$math_score,
          main = "Density Plot of Math Scores",
          xlab = "Math Scores")
```

Doesn't look very normal

Checking qqplot of Math Scores
```{r}
ggqqplot(exams$math_score)
```
Looks possibly normal

Shapiro-Wilks test
```{r}
shapiro.test(exams$math_score)
```
Since the p-value is below 0.05, the distribution is different from a normal distribution therefore we cannot assume the math scores are normally distributed


Let's see reading scores

```{r}
ggdensity(exams$reading_score,
          main = "Density Plot of Reading Scores",
          xlab = "Reading Scores")
```
```{r}
ggqqplot(exams$reading_score)
```
```{r}
shapiro.test(exams$reading_score)
```
Conclusion: reading score is not normal

```{r}
ggdensity(exams$writing_score,
          main = "Density Plot of Writing Scores",
          xlab = "Writing Scores")
```
```{r}
ggqqplot(exams$writing_score)
```
```{r}
shapiro.test(exams$writing_score)
```
Conclusion: Writing score is not normal

Linear models
seeing if variables (gender, race, pariental level of education, lunch, test prep score) are strongly related to scores

model <- lm(sales ~ youtube, data = marketing)
model

```{r}
model_math <- lm(data = exams,
            math_score ~ gender + race_ethnicity + parental_level_of_education + lunch + test_preparation_course)

summary(model_math)
```

can individually go through the residual and assumption plots with
plot(model_math, 1), plot(model_math,2), ....


## making plots for fun 


```{r boxplot of math score by gender}
ggplot(exams) + geom_boxplot(aes(y=math_score, x=race_ethnicity))
```


```{r histogram of math scores}
ggplot(exams) + geom_histogram(aes(math_score), bins=30)
```

plot3 <- ggplot(iris, aes(Sepal.Length, Petal.Length)) +
  geom_point()

```{r scatterplot of reading and writing score}
ggplot(exams) + geom_point(aes(x=reading_score, y=writing_score, color=gender)) +
  theme_classic()
```







