---
title: "analysis"
output: html_document
date: "2025-01-01"
---

website origin data 
https://www.kaggle.com/datasets/spscientist/students-performance-in-exams/code



Random notes and thoughts

Using if_else changes the TRUE/FALSE to numerical if T/F == 1/0
should literally put "0"/ "1". should change as.factor() afterwards




packages

```{r}
pacman::p_load(
  rio,          # File import
  here,         # File locator
  skimr,        # get overview of data
  tidyverse,    # data management + ggplot2 graphics, 
  gtsummary,    # summary statistics and tests
  rstatix,      # statistics
  corrr,        # correlation analayis for numeric variables
  janitor,      # adding totals and percents to tables
  flextable     # converting tables to HTML
)
```

importing data

```{r}
exams <- import("StudentsPerformance.csv")
```

```{r}
skim(exams)
```

Looks like we have 5 categorical columns and 3 continuous

Look like no missing data either :)

TO DO: clean the categorical data: make one-hot-encoding

First let's clean the column names

```{r}
exams <- exams %>% janitor::clean_names()
```

```{r}
colnames(exams)
```

Checking unique values before re-coding

```{r}
unique(exams$gender)
```

```{r}
exams$gender <- as.factor(if_else(exams$gender=="male",1,0))
```

checking values in race_ethnicity

```{r}
unique(exams$race_ethnicity)
```



trees.genus <- trees %>%
               mutate(Genus = case_when(               # creates the genus column and specifies conditions
                  grepl("Acer", LatinName) ~ "Acer",
                  grepl("Fraxinus", LatinName) ~ "Fraxinus"


creating race_ethnicity_f

```{r}
exams <- exams %>% 
  mutate(race_ethnicity_f = case_when(
    race_ethnicity == "group A" ~ 0,
    race_ethnicity == "group B" ~ 1,
    race_ethnicity == "group C" ~ 2,
    race_ethnicity == "group D" ~ 3,
    race_ethnicity == "group E" ~ 4,
  ))
```

changing into factor

```{r}
exams$race_ethnicity_f <- as.factor(exams$race_ethnicity_f)
```


```{r}
unique(exams$parental_level_of_education)
```


PERHAPS CHANGES THIS LOOK INTO R4EPI LESSON!


```{r}
exams$parental_level_of_education <- exams %>% 
  mutate(parental_level_of_education = case_when(
    parental_level_of_education == "some high school"   ~ 0,
    parental_level_of_education == "high school"        ~ 1,
    parental_level_of_education == "some college"       ~ 2,
    parental_level_of_education == "associate's degree" ~ 3,
    parental_level_of_education == "bachelor's degree"  ~ 4,
    parental_level_of_education == "master's degree"    ~ 5
  ))
```

exams



```{r}
unique(exams$lunch)
```

```{r}
unique(exams$test_preparation_course)
```











